Generalized Rybicki Press algorithm

### 2014-10-05 paper/ESS.tex
Fixing licenses and acknowledgement

The algorithm proposed in this article relies on two key ideas: (i) The covariance matrix in Equation~\eqref{eqn_covariance_matrix} is a semi-separable matrix with semi-separable rank $p$; (ii) Fast algorithms for solving and computing determinants for such semi-separable matrices can be obtained by embedding the dense covariance matrix in a large sparse banded matrix. The article also discusses how to perform the entire procedure in a numerically stable fashion. The implementation is made available at \url{https://github.com/sivaramambikasaran/ESS}~\cite{ambikasaran2014ESS} under the MIT license.

\textbf{Acknowledgements}: The author would like to thank Christopher S. Kochanek for initiating the conversation on generalized Rybicki Press algorithm and David W. Hogg for putting in touch with Christopher S. Kochanek. The author was supported in part by the Applied Mathematical Sciences Program of the U.S. Department of Energy under Contract DEFGO288ER25053, Office of the Assistant Secretary of Defense for Research and Engineering and AFOSR under NSSEFF Program Award FA9550-10-1-0180.

@@

%%%%%%%%%%%%%%%%%
%				%
%	CONCLUSION	%
%				%
%%%%%%%%%%%%%%%%%
\section{Conclusion}
The article discusses a numerically stable, generalized Rybicki Press algorithm, which relies on the fact that a semi-separable matrix can be embedded into a larger banded matrix. This enables $\mathcal{O}(N)$ inversion and determinant computation of covariance matrices, whose entries are sums of exponentials. This publication also serves to formally announce the release of the implementation of the extended sparse semi-separable factorization and the generalized Rybicki Press algorithm. The implementation is in C++ and is made available at \url{https://github.com/sivaramambikasaran/ESS}~\cite{ambikasaran2014ESS} under the MIT license.

>>>>

The algorithm proposed in this article relies on two key ideas: (i) The covariance matrix in Equation~\eqref{eqn_covariance_matrix} is a semi-separable matrix with semi-separable rank $p$; (ii) Fast algorithms for solving and computing determinants for such semi-separable matrices can be obtained by embedding the dense covariance matrix in a large sparse banded matrix. The article also discusses how to perform the entire procedure in a numerically stable fashion. The implementation is made available at \url{https://github.com/sivaramambikasaran/ESS}~\cite{ambikasaran2014ESS} under the license provided by New York University.

\textbf{Acknowledgements}: The author would like to thank Christopher S. Kochanek for initiating the conversation on generalized Rybicki Press algorithm and David W. Hogg for putting in touch with Christopher S. Kochanek. The research was supported in part by the NYU-AIG Partnership on Innovation for Global Resilience under grant number A2014-005. The author was also supported in part by the Applied Mathematical Sciences Program of the U.S. Department of Energy under Contract DEFGO288ER25053, Office of the Assistant Secretary of Defense for Research and Engineering and AFOSR under NSSEFF Program Award FA9550-10-1-0180.

@@

%%%%%%%%%%%%%%%%%
%				%
%	CONCLUSION	%
%				%
%%%%%%%%%%%%%%%%%
\section{Conclusion}
The article discusses a numerically stable, generalized Rybicki Press algorithm, which relies on the fact that a semi-separable matrix can be embedded into a larger banded matrix. This enables $\mathcal{O}(N)$ inversion and determinant computation of covariance matrices, whose entries are sums of exponentials. This publication also serves to formally announce the release of the implementation of the extended sparse semi-separable factorization and the generalized Rybicki Press algorithm. The implementation is in C++ and is made available at \url{https://github.com/sivaramambikasaran/ESS}~\cite{ambikasaran2014ESS} under the license provided by New York University.
